---
title: "CircSpaceTime: An R package for spatial, spatio-temporal and temporal model for circular, cylindrical and spherical data"
author: |
  |
  | <a href="https://github.com/santoroma/CircSpaceTime">github.com/santoroma/CircSpaceTime</a></small>
  | <small>23rd International Conference on Computational Statistics</small>
output:
  revealjs::revealjs_presentation:
    template: COMPSTAT_2018_Template.html
    incremental: false
    self_contained: false
    reveal_plugins: ["notes", "zoom", "menu"]
    theme: league
    transition: zoom
    background_transition: zoom
    highlight: pygments
    center: true
    progress: true
    slideNumber: true
    menu:
        numbers: true
    defaultTiming: 120
---

## A brand new package {data-background="img/storm1.gif"}


- Currently the following models are implemented:  
  
    + Spatial Wrapped Normal
    + Spatial Projected Normal
<br>
- Yet to come (in few weeks):  

    + Spatio-Temporal Wrapped Normal
    + Spatio-Temporal Projected Normal  
<br>


## A brand new package {data-background="img/inception_end.gif"}

- Already available on GitHub at <a href="https://github.com/santoroma/CircSpaceTime">github.com/santoroma/CircSpaceTime</a>  
The package will be released on CRAN before the end of the 2018.
<br>

- It will be constantly updated with Bayesian and classical models dealing with complex dependence structures for circular, cylindrical and spherical variable. 
<br>  
   


# A brief example  {data-background="img/map1.png"}


##   {data-background="img/map1.png"}  



- <span style="background-color:black"> Example based on wave directions (and heights): a storm event observed at 8pm of April 6, 2010.</span> 
<br>    
<br>

- <span style="background-color:black"> Data available inside the package.</span>
<br>

##  Test data {data-background="img/plot2.png" data-background-size="100%" data-background-position="bottom"}  

 We hold out 10% of the locations for validation purposes 
<br>    
<br><br>
<br><br><br>
<br>    
<br><br>
<br><br><br>
<br><br><br>


# Estimation based on the Wrapped Gaussian


##  WrapSp   
  
<small>  

- **WrapSp** function produces samples from the Wrapped Normal spatial model posterior distribution  
- As inputs it requires:  

    + a vector of $n$ circular data in $[0,2\pi)$ and a matrix of coordinates   (*train data* in our example)
    + two lists, one of starting values for the MCMC and the other for the prior distributions definition.
    + Further inputs are related to computational options such as the number of chains, if the computation should be parallelized and the number of iterations.  


```{r wrapsp, warning = FALSE, echo=TRUE, eval=FALSE, size="tiny"}
 storm <- WrapSp(
 x     = train0$Dmr,
 coords    = coords0.train,
 start   = start0 ,
 prior   = list("alpha"      = c(pi,10), 
 "rho"     = c(rho_min0, rho_max0), 
 "sigma2"    = c(3,0.5),
 "beta"      = c(1,1,2)  
 ) ,
 nugget = TRUE,
 sd_prop   = list( "sigma2" = 1, "rho" = 0.3, "beta" = 1),
 iter    = 30000,
  bigSim    = c(burnin = 15000, thin = 10),
 accept_ratio = 0.5,
 adapt_param = c(start = 1000, end = 10000, esponente = 0.95),
 corr_fun = "exponential", 
 n_chains = 2,
 parallel = T,
 n_cores = 2)
```
</small>

##  WrapKrig  
  
<small>  

- **WrapKrig** function estimate the values on the test sites using the posterior samples we just obtained 
- As inputs it requires:  

    + the output of *WrapSp*   
    + the coordinates for the train (observed) points
    + the coordinates of the test (validation) points
    + the observed (train) circular values

</small>
```{r wrapkrig, warning = FALSE, echo=TRUE, eval=FALSE}
 Pred.storm <- WrapKrig(
   WrapSp_out = storm,
## The coordinates for the observed points
  coords_obs = coords0.train,
## The coordinates of the validation points
  coords_nobs = coords0.test,
##the observed circular values
   x_oss = train0$Dmr
 )
```

# Estimation based on the Projected Gaussian


##  ProjSp   
  
<small>  

- **ProjSp** function produces samples from the Projected Normal spatial model posterior distribution  
- As inputs it requires:  

    + a vector of $n$ circular data in $[0,2\pi)$ and a matrix of coordinates   (*train data* in our example)
    + two lists, one of starting values for the MCMC and the other for the prior distributions definition.
    + Further inputs are related to computational options such as the number of chains, if the computation should be parallelized and the number of iterations.  


```{r projsp, warning = FALSE, echo=TRUE, eval=FALSE, size="tiny"}
 mod0_PN <- ProjSp(
  x     = train0$Dmr,
  coords    = coords0.train,
  start   = start0_PN ,
  prior   = list("alpha_mu"      = c(0,0),
                 "alpha_sigma"   = diag(10,2),
                 "rho0"     = c(rho_min0, rho_max0),
                 "rho"      = c(-1,1),
                 "sigma2"    = c(3,0.5)),
  sd_prop   = list( "sigma2" = .1, "rho0" = 0.1, "rho" = .1,  "sdr" = sample(.05,length(train0$Dmr), replace = T)),
  iter    = 5000,
  bigSim    = c(burnin = 3500, thin = 1),
  accept_ratio = 0.5,
  adapt_param = c(start = 1000, end = 10000, esponente = 0.95, sdr_update_iter = 50),
  corr_fun = "exponential", 
  n_chains = 2,
  parallel = T,
  n_cores = 2)
```
</small>

##  ProjKrig  
  
<small>  

- **ProjKrig** function estimate the values on the test sites using the posterior samples we just obtained 
- As inputs it requires:  

    + the output of *ProjSp*   
    + the coordinates for the train (observed) points
    + the coordinates of the test (validation) points
    + the observed (train) circular values

</small>
```{r projkrig, warning = FALSE, echo=TRUE, eval=FALSE}
 Pred.krig_PN <- ProjKrig(mod0_PN,
                      ## The coordinates for the observed points  
                      coords_obs = coords0.train,
                      ## The coordinates of the validation points
                      coords_nobs = coords0.test,
                      ##the observed circular values
                      x_oss = train0$Dmr)
```


## Comparison

- We can compare the predictions of the two models using the Average Prediction Error (APE)
<br>
- It's the package function **APEcirc**
<br>
<br>
<br>
```{r print_APE_comparison, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- " 
|                           | Wrapped | Projected |
|---------------------------|:-------:|----------:|
| Average Prediction Error  | 0.0007  | 0.0010    |
"
cat(tabl) 
```


```{r ape, warning = FALSE, echo=TRUE, eval=FALSE, size="tiny"}
 APE_WRAP <- APEcirc( real = test0$Dmr,
                sim = Pred.storm$Prev_out,
                bycol = F
)
  APE_PN <- APEcirc( real = test0$Dmr,
                sim = Pred.krig_PN$Prev_out,
                bycol = F
)
```


# ... this is the END {data-background="img/truman_show_end.gif"}

## THANKS!!! {data-background="img/truman_show_end.gif"}

Further information and installation instructions on
<font size="+14">[github.com/santoroma/CircSpaceTime](https://github.com/santoroma/CircSpaceTime)</font>

